### 2025-04-04 documentation/hierdata.qmd

---
title: "hierarchical data"
author: "leontine"
format: html
editor: visual
---

Load relevant packages, including our own and data

```{r}
#| include: false
library(here) 
library(tidyverse) 
library(ggplot2) 
library(haven) 
library(stringr)
library(cmdstanr) 
library(bayesplot) # diagnostic plots 
devtools::load_all(here::here())
#data_folder <- "data_raw" 
# calling survey_df data here, as internal to fit_model
#data <- read_csv(here::here(data_folder, "coverage_data.csv"))
```

toy data set

```{r}
data <- tibble(subcluster = c("A", "A", "B", "B"), iso = c("iso1", "iso2", "iso3", "iso3"))
```

main info: hierarchical levels to use, and the level of interest

(level of interest does not need to be the lowest level, not clear how thoroughly tested)

```{r}
hierarchical_level <- c("intercept",  "subcluster", "iso")
area <- "iso"
```

Create district index for matching district and district index

```{r}
geo_unit_index <- get_geo_unit_index_data(data,
                                          hierarchical_levels = c(hierarchical_level), 
                                          area = area)
geo_unit_index
```

Set up hierarchical structures

```{r}
# use in fit_model:
#hier_data <-  list()
#hier_data[["mu_data"]] <- 
# here we just work with one object:
hier_data <- hierarchical_data(geo_unit_index, hierarchical_level)
#names(hier_data[["mu_data"]])
names(hier_data)
hier_data

```

for use in stan

```{r}
hier_stan_data <- hierarchical_param_stan_data(
  global_fit = NULL,
  param_name ="mu",
  param_data = hier_data #[["mu_data"]] #,
  # used only when fixing things 
#  hierarchical_terms_fixed = hierarchical_level_terms_fixed,
#  hierarchical_sigmas_fixed = hierarchical_level_sigmas_fixed
)
names(hier_stan_data)
```

```{r}
hier_stan_data$mu_n_terms # number of etas + global mu
length(unique(geo_unit_index$subcluster))
length(unique(geo_unit_index$iso))

```

```{r}
hier_stan_data$mu_n_sigma # update: no longer includes prior sd for intercept 
```

Let's illustrate how these objects are used in R and in the stan model.

For that, make up some values for the parameters based on

```         
"mu_n_terms"          "mu_n_sigma"
```

```{r}
#mu_raw refers to the vector with ALL $\gamma$s, including the global one.The ordering is global-allregional-allsubregional-country.
mu_raw <- seq(1, hier_stan_data$mu_n_terms)
mu_raw
#mu_sigma referred to the vector of stacked sigmas, excluding $s_{\mu, global}$ (in the #original package, it was called mu_sigma_raw).
mu_sigma <- seq(1, hier_stan_data$mu_n_sigma)
sprior <- 0.5

```

try to expose stan functions using cmdstanr (not rstan)

<https://mc-stan.org/cmdstanr/reference/model-method-expose_functions.html>

```{r}
stan_file <- here("inst/stan/hierfunctions_seq.stan")
mod <- cmdstan_model(stan_file, force_recompile = TRUE)  
mod$expose_functions(global = TRUE) # to add to global environment

```

#mu_star refers to the vector with (mu global and ALL $\eta$s). we get mu_stars from mu_raw \# and the vector ($s_{\mu, global}$ , mu_sigma), using the stan scale_blocks function.

```{r}
mu_star <- scale_blocks(mu_raw, c(sprior, mu_sigma), 
             start = hier_stan_data$mu_re_start,
             end = hier_stan_data$mu_re_end)
mu_star
```

mu refers to the final $\mu$s at the country level. Matrix multiplication is used to obtain the mu country vector from the mu_stars. The full matrix mu_model_matrix is a sparse matrix, with 1s and 0s (columns refer to $\mu_{global}$, the regional etas, the subregional etas, the country etas). The matrix multiplication is implemented using the stan function csr_matrix_times_vector, using a decomposition of mu_model_matrix into different parts U, V, W.

check if this is right

```{r}
mu <- hier_stan_data$mu_model_matrix%*%mu_star
mu
```

#

```{r}
hier_data$model_matrix
```

# Unsuccessful investigation into exposing stan build-in functions

Need to expose (compile) the stan math functions as well, eg for for sparse matrices. see here for info

<https://mc-stan.org/docs/functions-reference/sparse_matrix_operations.html#conversion-functions>

<https://stan-dev.r-universe.dev/articles/StanHeaders/stanmath.html>

<https://discourse.mc-stan.org/t/expose-stan-functions-with-stan-math-library-function/15771/6>

help(stanFunction, package = "StanHeaders")

first time around, i got an error

Error in sourceCpp(code = code, env = env, rebuild = rebuild, cacheDir = cacheDir, : Error 1 occurred building shared library.

Turns out that it may not work anyway:

looks like csr_extract_w would be tricky... (was getting errors anyway with something along lines of StanHeaders::stanFunction("csr_extract_w"))

<https://discourse.mc-stan.org/t/function-to-return-number-of-non-zero-elements-in-a-matrix/26123/7>

looks like we just want model_matrix\*vector

and can confirm if

csr_matrix_times_vector

```{r}
#csr_extract_w()
```

```{r}
# get_onehierparam(oneparam_sigma = c(sprior, mu_sigma),  
#                  oneparam_raw_fixed = mu_raw[1], #numeric(0), 
#                  oneparam_raw_estimate = mu_raw[-1],
#                  oneparam_re_start = hier_stan_data$mu_re_start,
#                  oneparam_re_end = hier_stan_data$mu_re_end, 
#                  C = nrow(geo_unit_index), 
#                  oneparam_n_terms = hier_stan_data$mu_n_terms, 
#                  oneparam_n_terms_fixed = 1, #hier_stan_data$mu_n_terms_fixed, 
#                  oneparam_n_terms_estimate = hier_stan_data$mu_n_terms_estimate -1,
#       oneparam_model_matrix_w= csr_extract_w(hier_stan_data$mu_model_matrix), 
#       oneparam_model_matrix_v = hier_stan_data$mu_model_matrix_v, 
#       oneparam_model_matrix_u = hier_stan_data$mu_model_matrix_u
#           )
```