### 2025-09-30 paper/title.tex
misc paper editing, refs

\title{localhierarchy: an R package to facilitate fitting of global and local Bayesian hierarchical models}

\author[1]{Leontine Alkema\thanks{
This work was supported, in whole or in part, by the Bill \& Melinda Gates Foundation (INV-00844 and OPP1192802).
Under the grant conditions of the Foundation, a Creative Commons Attribution 4.0 Generic License has already been assigned to the Author Accepted Manuscript version that might arise from this submission.
Contact: lalkema@umass.edu.}}
\affil[1]{\textit{Department of Biostatistics and Epidemiology, %School of Public Health and Health Sciences,
University of Massachusetts Amherst, %Amherst, %Massachusetts,
USA}}

\author[2]{Shauna Mooney}
\affil[2]{\textit{Hamilton Institute and Department of Mathematics and Statistics, Maynooth University, Ireland}}

\author[1]{Evan Ray}

\author[3]{Herbert Susmann}
\affil[3]{\textit{Division of Biostatistics, Department of Population Health, NYU, % Grossman School of Medicine, %New York, %New York,
USA}}
\date{\today}
\maketitle