### 2025-09-30 README.Rmd
cleaning

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# localHierarchy

The `localHierarchy` R package provides functionality for fitting Bayesian hierarchical models in settings where both global and local estimation is required. The package provides R functions and Stan model components to support global modeling, in which all parameters in hierarchical models are estimated, and local modeling, where parameters are estimated for only one or a small number of populations, using fixed values from a global model fit.

## Installation

-   Dependency on `cmdstanr:` Instructions for installing `cmdstanr` are available in their [Getting started](https://mc-stan.org/cmdstanr/articles/cmdstanr.html) guide.

-   Install `localHierarchy` from Github: `remotes::install_github("AlkemaLab/localhierarchy")`

## Examples

to do: add from paper