### 2025-11-21 README.Rmd
update readme

---
title: "localhierarchy"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

## Background

Bayesian hierarchical models are widely used in global health estimation, where data availability may vary across national or subnational populations. Such models are typically fitted to a global database, e.g., to produce national-level estimates for all countries in the world or in some region. To facilitate analysis at a local level, models are often desirable that are informed by global models but can be fitted to just a subset of the data, such as data for one country. We refer to such models as local models: models that are derived from global hierarchical models but adapted such that they can be fitted to the data from one population alone.

The `localhierarchy` R package provides functionality for fitting Bayesian hierarchical models in settings where both global and local estimation is required. The package provides R functions and Stan model components to support global modeling, in which all parameters in hierarchical models are estimated, and local modeling, in which parameters are estimated for only one or a small number of populations, using fixed values from a global model fit.

The article on this website presents a practical introduction to the package for the applied user and illustrates the package’s functionality through examples for national estimation.

## Installation

-   `localhierarchy` depends on `cmdstanr:` Instructions for installing `cmdstanr` are available in their [Getting started](https://mc-stan.org/cmdstanr/articles/cmdstanr.html) guide.

-   Install `localhierarchy` from Github: `remotes::install_github("AlkemaLab/localhierarchy")`

## Examples

See [article](https://alkemalab.github.io/localhierarchy/articles/localhierarchy_article.html) on package website, at [https://alkemalab.github.io/localhierarchy](https://alkemalab.github.io/localhierarchy/index.html){.uri}.

## Citation

Please cite as follows:

```{r}
citation("localhierarchy")
```